---
title: "Class 9 Halloween Project"
author: "Liana Melikian (A16675734)"
format: pdf
---

```{r}
candy_file="candy-data.csv"
candy=read.csv(candy_file,row.names=1)
head(candy)
```
Q1
```{r}
nrow(candy)
```
Q2
```{r}
sum(candy$fruity)
```

Q3
```{r}
candy["Snickers",]$winpercent
```

Q4
```{r}
candy["Kit Kat",]$winpercent
```

Q5
```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
library(skimr)
skim(candy)
```

Q6
winpercent

Q7
1 means the candy is made of chocolate, and 0 means it isn't.

Q8
```{r}
hist(candy$winpercent)
```
Q9.
No, it is skewed to the right.

Q10. Above
```{r}
mean(candy$winpercent)
```

Q11. Higher
```{r}
choc.inds=as.logical(candy$chocolate)
choc.win=candy[choc.inds,"winpercent"]
mean(choc.win)
fruit.inds=as.logical(candy$fruity)
fruit.win=candy[fruit.inds,"winpercent"]
mean(fruit.win)
mean(choc.win)>mean(fruit.win)
```

Q12. Yes, p<0.05
```{r}
t_test_result=t.test(choc.win, fruit.win)
print(t_test_result)
```

Q13. Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters
```{r}
head(candy[order(candy$winpercent),], n=5)
```
Q14. Reese's PB cup, Reese's miniatures, Twix, Kit Kat, Snickers
```{r}
head(candy[order(-candy$winpercent), ], n = 5)
```
Q15.
```{r}
library(ggplot2)
ggplot(candy)+
  aes(winpercent, rownames(candy))+
  geom_col(fill="blue")
```
Q16.
```{r}
ggplot(candy)+
  aes(winpercent, reorder(rownames(candy),winpercent))+
  geom_col(fill="blue")
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
 
 Q17. Sixlets
 
 Q18. Starburst
 
```{r}
library(ggrepel)
# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
Q19. Reeses miniatures

Q20. Nik L Nip, Nestle Smarties, Ring pop, Hersheys Krackel, Hersheys Milk Chocolate
Nik L Nip is least popular
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

Q22. chocolate and fruity
Q23. chocolate and winpercent

```{r}
pca= prcomp(candy, scale = TRUE)
summary(pca)
```
```{r}
plot(pca$x[, 1], pca$x[, 2], 
     xlab = "PC1", ylab = "PC2")
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
#ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24. Fruity, hard, and pluribus. Yes these make sense to me.
