---
title: "Class05: Data Visualization with GGPLOT"
author: "Liana Melikian (ID: A16675734)"
format: pdf
---

## Quarto

---
title: "BIMM 143: Class 5 - Data Visualization with ggplot2"
author: "Liana Melikian - A16675734"
format: pdf
---

## Using GGPLOT

The ggplot 2 package needs to be installed as it does not come with R "out of the box."

We use the `install.packages()` function to do this.

```{r}
head(cars)
```

To use ggplot I need to load it up before I can call any of the functions in the package. I do this with the `library()` function.

```{r}
library(ggplot2)
```
All ggplot figures have at least 3 things:
-data (the stuff we want to plot)
-aesthetic mapping (aes vales)
-geoms

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point()
```


```{r}
ggplot(cars)+
  aes(x=speed,y=dist)+
  geom_point()+
  labs(title="Speed and Stopping Distances of Cars", x="Speed (MPH)", y="Stopping Distance (ft)", subtitle = "Your informative subtitle text here", caption = "Dataset:'cars'")+
  geom_smooth(method="lm",se=FALSE)+
  theme_bw()
```

ggplot is not the only graphing system in R, there are lots of others. There is even "base R" graphics.

```{r}
plot(cars)
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
colnames(genes)
ncol(genes)
table(genes$State)
round(table(genes$State)/nrow(genes)*100,2)
```
```{r}
ggplot(genes) + 
    aes(x=Condition1, y=Condition2) +
    geom_point()
```
```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
p
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red") )
```

```{r}
p=ggplot(genes)+
  aes(x=Condition1,y=Condition2,col=State)+
  geom_point()
p+scale_color_manual(values=c("blue","gray","red"))+
  labs(title="Gene Expression Changes Upon Drug Treatment",x="Control (no drug)",y="Drug Treatment")
```
```{r}
library(gapminder)
data("gapminder")
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```
```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
```
```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
```
```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5)
```

```{r}
# install.packages("dplyr")
library(dplyr)
```
```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
library(ggplot2)
ggplot(gapminder_2007)+
  geom_point(aes(x=gdpPercap, y=lifeExp,size=pop),alpha=0.5)+
  scale_size_area(max_size=10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)
ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```
```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop))
```

```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = continent))
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = lifeExp))
```
```{r}
ggplot(gapminder_top5) +
  aes(x=country, y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=gdpPercap) +
  geom_col()
```

```{r}
ggplot(gapminder_top5)+
  aes(x=reorder(country,-pop),y=pop, fill=country)+
  geom_col(col="gray30")+
  guides(fill="none")
```
```{r}
head(USArrests)
USArrests$State=rownames(USArrests)
ggplot(USArrests)+
  aes(x=reorder(State,Murder),y=Murder)+
  geom_col()+
  coord_flip()
```
```{r}
ggplot(USArrests)+
  aes(x=reorder(State,Murder),y=Murder)+
  geom_point()+
  geom_segment(aes(x=State,xend=State,y=0, yend=Murder), color="blue")+
  coord_flip()
```
```{r}
#install.packages("gifski") 
#install.packages("gganimate")
library(gapminder)
library(gganimate)
```
```{r}
# Setup nice regular ggplot of the gapminder data
#ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  #geom_point(alpha = 0.7, show.legend = FALSE) +
  #scale_colour_manual(values = country_colors) +
  #scale_size(range = c(2, 12)) +
  #scale_x_log10() +
  # Facet by continent
  #facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  #labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  #transition_time(year) +
 # shadow_wake(wake_length = 0.1, alpha = FALSE)
```

```{r}
library(patchwork)

p1 <- ggplot(mtcars) + geom_point(aes(mpg, disp))
p2 <- ggplot(mtcars) + geom_boxplot(aes(gear, disp, group = gear))
p3 <- ggplot(mtcars) + geom_smooth(aes(disp, qsec))
p4 <- ggplot(mtcars) + geom_bar(aes(carb))
(p1 | p2 | p3) /
      p4
```

```{r}
library(ggplot2)
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  geom_smooth()
```

```{r}
library(ggplot2)
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  geom_smooth(method="lm",se=FALSE)
```

```{r}
library(ggplot2)
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  labs(title="Speed and Stopping Distance of Cars",x="Speed (MPH)",y="Stopping Distance(ft)",subtitle="LianaloveAra",caption="Dataset: 'cars'")+
  geom_smooth(method="lm",se=FALSE)+
  theme_bw()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
ncol(genes)
colnames(genes)
```

```{r}
table(genes$State)
table(genes$State)["up"]

```
```{r}
round(table(genes$State)/nrow(genes)*100,2)
```

```{r}
p=ggplot(genes)+
  aes(x=Condition1,y=Condition2,col=State)+
  geom_point()
p+scale_color_manual(values=c("blue","gray","red"))+ 
  labs(title="Gene Expression Changes Upon Drug Treatment",x="Control (no drug)",y="Drug Treatment")
```

```{r}
library(gapminder)
library(dplyr)
gapminder_2007 <- gapminder %>% filter(year==2007)
```
```{r}
head(gapminder_2007)
```

```{r}
ggplot(gapminder_2007)+
  aes(x=gdpPercap,y=lifeExp,size=pop)+
  geom_point(alpha=0.5)+
  scale_size_area(max_size=10)
```

